Predicting the 2024 U.S. Presidential Election: Trump’s Likely Victory Based on Polling Analysis

### 2024-11-03 paper/paper.qmd
commit changes to paper.qmd in results section

## National Support Trends Over Time

```{r}
#| echo: false
#| eval: true
#| label: fig-national-trend
#| fig-cap: "National Support Trends Over Time for the 2024 U.S. Presidential Election. The plot displays the predicted national average support for Donald Trump (red) and Kamala Harris (blue) over time, based on a Bayesian hierarchical model with a natural spline fit. The trends show changes in polling support for both candidates throughout 2024, highlighting key shifts in public opinion as the election approaches."
#| warning: false

# Generate a sequence of dates for national predictions
prediction_dates <- seq.Date(min(clean_poll_data$end_date), max(clean_poll_data$end_date), by = "1 week")

# Predict national support for each date
national_trend <- prediction_dates %>%
  map_dfr(~ {
    data <- data.frame(state = unique(clean_poll_data$state), end_date = .x)
    data$end_date_num <- as.numeric(data$end_date - min(clean_poll_data$end_date))

    trump_pred <- mean(posterior_predict(trump_model, newdata = data))
    harris_pred <- mean(posterior_predict(harris_model, newdata = data))

    tibble(date = .x, Trump = trump_pred, Harris = harris_pred)
  })

# Plot national support trends
ggplot(national_trend, aes(x = date)) +
  geom_line(aes(y = Trump, color = "Trump")) +
  geom_line(aes(y = Harris, color = "Harris")) +
  labs(title = "National Support Trends Over Time", y = "Predicted Support", x = "Date") +
  scale_color_manual(values = c("Trump" = "red", "Harris" = "blue")) +
  theme_minimal()

```
EDIT - explain results on graph

## State Level Support Trends Over Time
```{r}
#| echo: false
#| eval: true
#| label: fig-state-trend
#| fig-cap: "State-Level Support Trends for Trump and Harris Over Time."
#| warning: false

@@

# Generate predictions for each state and date
state_dates <- seq.Date(min(clean_poll_data$end_date), max(clean_poll_data$end_date), by = "1 week")
state_prediction_grid <- expand.grid(state = battleground_states, end_date = state_dates) %>%
  mutate(end_date_num = as.numeric(end_date - min(clean_poll_data$end_date)))

# Predict support for each state and date
state_predictions <- state_prediction_grid %>%
  rowwise() %>%
  mutate(
    Trump_Support = mean(posterior_predict(trump_model, newdata = cur_data())),
    Harris_Support = mean(posterior_predict(harris_model, newdata = cur_data()))
  )

# Plot state-level support trends over time with adjusted x-axis labels
ggplot(state_predictions, aes(x = end_date)) +
  geom_line(aes(y = Trump_Support, color = "Trump")) +
  geom_line(aes(y = Harris_Support, color = "Harris")) +
  labs(title = "State-Level Support Trends for Trump and Harris Over Time (Battleground States)", 
       y = "Predicted Support", 
       x = "Date") +
  facet_wrap(~ state) +
  scale_color_manual(values = c("Trump" = "red", "Harris" = "blue")) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +  # Set date format and break frequency
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    strip.text = element_text(size = 10),  # Control facet label size if needed
    panel.spacing = unit(0.5, "lines"),    # Add spacing between facets
    panel.grid.major = element_line(linewidth = 0.3)  # Adjust gridline thickness
  )

```
EDIT - explain results on graph

>>>>

## State Level Support Trends Over Time
```{r}
#| echo: false
#| eval: true
#| label: fig-state-polling-average
#| fig-cap: "Polling Averages for Trump and Harris Over Time in Battleground States."
#| warning: false

@@

# Calculate average polling percentages over time for battleground states
polling_average <- clean_poll_data %>%
  filter(state %in% battleground_states) %>%
  group_by(end_date, state) %>%
  summarise(
    Average_Trump = mean(Trump_pct, na.rm = TRUE),
    Average_Harris = mean(Harris_pct, na.rm = TRUE),
    .groups = 'drop'
  )

# Plot polling averages over time for each battleground state
ggplot(polling_average, aes(x = end_date)) +
  geom_line(aes(y = Average_Trump, color = "Trump")) +
  geom_line(aes(y = Average_Harris, color = "Harris")) +
  labs(title = "Polling Averages for Trump and Harris Over Time in Battleground States", 
       y = "Average Support (%)", 
       x = "Date") +
  facet_wrap(~ state) +
  scale_color_manual(values = c("Trump" = "red", "Harris" = "blue")) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +  # Set date format and break frequency
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    strip.text = element_text(size = 10),  # Control facet label size if needed
    panel.spacing = unit(0.5, "lines"),    # Add spacing between facets
    panel.grid.major = element_line(linewidth = 0.3)  # Adjust gridline thickness
  )

EDIT - explain results on graph
